---
title: "Exploratory Analysis of Covid-19 Data"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    horizontal_layout: fill
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(naniar)
library(plotly)
library(rvest)
library(ggplot2)
library(lubridate)

library(flexdashboard)
```

Row {data-height=650}
-----------------------------------------------------------------------

### Covid 19 Case Rate Per 100k

```{r, message=FALSE, warning=FALSE}
# input data table for Total Cases by State/Territory
covid_total_by_state = 
  read_csv("./data/united_states_covid19_cases_deaths_and_testing_by_state.csv", skip = 2) %>% 
  janitor::clean_names()
# figure out the missing data in data frame
na_strings = "N/A"
# fill in missing data
covid_total_by_state_tidy =
  covid_total_by_state %>% 
  replace_with_na_all(condition = ~.x %in% na_strings) %>% 
  mutate(
    state_territory = as.factor(state_territory),
    case_rate_2021 = as.numeric(case_rate_per_100000)
         ) %>% 
  rename(state = "state_territory") %>% 
  select(state, case_rate_2021) %>% 
  mutate(state = recode(state, "New York*" = "New York"))

#Let's make a choropleth map plot to show the total covid cases by state in US.
#Load the usa postal code dataset.
postal_code = 
  read_csv("./data/us_postal_code.csv") %>% 
  janitor::clean_names()

# read data function
read_data_function = function(file_df) {
  
  state_df = 
    read_csv(file_df, skip = 2)
  
  return(state_df)
}
# create a data frame containing all participants
state_2020 =
  tibble(
    files = list.files("./data/state_covid_separate")
  ) %>% 
  mutate(
    path = map(.x = files, ~paste("./data/state_covid_separate", ., sep = "/"))
  ) %>%                                                             # add path
  mutate(
    observations = map(path, read_data_function))

state_2020_tidy =
  state_2020 %>% 
  unnest(cols = "observations") %>% 
  janitor::clean_names() %>% 
  filter(date == "Dec 31 2020") %>%
  select(state, total_cases, total_case_rate_per_100k)

plot_df = left_join(state_2020_tidy, covid_total_by_state_tidy, postal_code, by = "state") %>% 
  left_join(postal_code, by = "state") %>% 
  mutate(plot_text = str_c(state, total_case_rate_per_100k, sep = ":"),
         plot_text2 = str_c(state, case_rate_2021, sep = ":"))

g = list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

plot_df$hover = with(plot_df, paste(
  state, '<br>', "Total case", total_cases, "<br>")
  )

#Make map plot.
map_plotly = 
  plot_geo(plot_df, locationmode = 'USA-states') %>% 
  add_trace(
    type = "scattergeo",
    locations = ~code,
    text = ~code,
    mode = "text",
    textfont = list(color = rgb(0,0,0), size = 8),
    hoverinfo = "none"
  ) %>% 
  add_trace(
    z = ~total_case_rate_per_100k, 
    text = ~plot_text, 
    locations = ~code,
    color = ~total_case_rate_per_100k, 
    colorscale = list(c(0,0.5,1), c("#ffffff", "#83c5fd", "#0066FF")),
    colorbar = list(title = "Case Rate"),
    hoverinfo = "text",
    showlegend = FALSE,
    visible = T
) %>% 
  add_trace(
    z = ~case_rate_2021, 
    text = ~plot_text2, 
    locations = ~code,
    color = ~case_rate_2021, 
    colorscale = list(c(0,0.6,1), c("#ffffff", "#83c5fd", "#0066FF")),
    colorbar = list(title = "Case Rate"),
    hoverinfo = "text",
    showlegend = FALSE,
    visible = F
  ) %>% 
  layout(
    title = "Case Rate of Covid 19 Per 100k population of all States",
    heights = 600,
    geo = g,
    updatemenus = list(
      list(
        type = 'buttons',
        x = 0.1,
        y = 0.95,
        buttons = list(
          list(method = "restyle",
               args = list("visible", list(T, T, F)),
               label = '2020'),
          
          list(method = "restyle",
               args = list("visible", list(T, F, T)),
               label = '2021')
        )))
  )

map_plotly
```

### Trends in Number of Cases

```{r, message=FALSE, warning=FALSE}
covid_cum = read_csv("data/covid_cumulative_cases.csv", skip = 2) %>% 
  janitor::clean_names()
covid_day = read_csv("data/covid_daily_cases.csv", skip = 2) %>% 
  janitor::clean_names()
covid_daily = left_join(covid_day, covid_cum, by = "date") %>% 
  select(date, new_cases, total_cases) %>% 
  separate(date, into = c("month","day","year")) %>% 
  mutate(month = factor(month, levels = month.abb)) %>% 
  group_by(year, month) %>% 
  arrange(year, month, day) %>% 
  mutate(date = make_date(year, month, day)) %>% 
  arrange(date)

row_cum = nrow(covid_cum)
row_daily = nrow(covid_daily)
daily_fig = plot_ly(covid_daily) 
daily_fig %>% 
  add_trace(x = ~date, y = ~new_cases, type = "bar", yaxis="y", name = "new", marker = list(color = 'rgb(158,202,225)')) %>% 
  add_trace(x = ~date, y = ~total_cases, type = "scatter", mode = "lines", yaxis = "y2", name = "cum") %>% 
  layout(yaxis=list(title = "daily new cases", side="left"),
         yaxis2=list(title = "cumulative cases", side="right",overlaying="y"),
         showlegend=TRUE)
```

Row {data-height=350}
-----------------------------------------------------------------------

### Comments
The dataset of Covid 19 in the United States begins from January 17, 2020. Considering the differences in population of each state, the map displays the total number of cases per 100k people in all states of the US in 2020 and 2021.

The first peak of number of positive cases in the US occurred around April 2020. The number goes up and down and reached to the highest value at the beginning of 2021. The cumulative number of cases keep rising ever since the pandemic started.

Currently California has the most number of positive cases among all states. If analyze based on population, North Dakota has the largest case rate in 2020 and 2021. From the map of case rate, we can see states from the central part of US tend to have larger value of case rate although the total number of cases are not higher than some other states with larger population. 

### Total Number of cases by state

```{r}
plot_df %>% 
  mutate(abbrev = fct_reorder(as.factor(abbrev), total_cases)) %>% 
  plot_ly(x = ~abbrev, y = ~total_cases, color = ~abbrev, type = "bar", colors = "viridis") %>% 
  layout(yaxis = list(title = "Total Cases"),
         xaxis = list(title = "State"),
         showlegend = FALSE)
```

