---
title: "The association between covid and state in US"
author: "Yujin Zhang"
output: github_document
---

This is an R markdown document about covid vs. state data analysis

```{r}
library(tidyverse)
library(naniar)
library(usmap)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

## Part 1
In this part I input the CDC covid cases data by state with time period till Thu Nov 18 2021.

The data generated: Thu Nov 18 2021 22:02:06 GMT-0500 (EST)

```{r}
# input data table for Total Cases by State/Territory
covid_total_by_state = 
  read_csv("./data/united_states_covid19_cases_deaths_and_testing_by_state.csv", skip = 2) %>% 
  janitor::clean_names()
```

For the missing data, the original dataset fill it with "N/A", so I will replace it with NA and change the variable to numeric.

```{r}
# figure out the missing data in data frame
na_strings = "N/A"

# fill in missing data
covid_total_by_state_tidy =
  covid_total_by_state %>% 
  replace_with_na_all(condition = ~.x %in% na_strings) %>% 
  mutate(
    state_territory = as.factor(state_territory),
    case_rate_per_100000 = as.numeric(case_rate_per_100000)
         ) %>% 
  rename(region = "state_territory") %>% 
  select(region, case_rate_per_100000)
  
```

In the following steps, I will make a choropleth map plot to show the total covid cases by state in US.

First create a new dataset for plot.

```{r}
# load geographic coordinates of each US state
all_state = usmap::us_map() %>% 
  relocate(full)

# combine two dataset
plot_df = left_join(covid_total_by_state_tidy, all_state, by = c("region" = "full"))
```

Add state abbreviations to the map figure.

```{r}
us_centroids = 
  plot_df %>% 
  group_by(region) %>% 
  summarise(centroid.x = mean(range(x)), 
            centroid.y = mean(range(y)),
            label = unique(toupper(str_sub(region,1,2))),
            case_rate_per_100000 = unique(case_rate_per_100000))

```

Now, let's draw the plot.

```{r}
plot_df %>% 
  group_by(region) %>% 
  ggplot() +
  geom_polygon(aes(x, y, group = group, fill = case_rate_per_100000)) +
  geom_text(
    data = us_centroids, 
    aes(centroid.x, centroid.y, label = paste(label, "\n", case_rate_per_100000)),
    size = 1
  ) +
  labs(title = "COVID-19 Case Rate by State/Territory (cases per 100,000)", fill = "Cases Rate Per 100,000") +
  # hide ticks on x and y axis
  scale_y_continuous(breaks = c()) + 
  scale_x_continuous(breaks = c()) +
  scale_fill_gradient2(low = "white", mid = "skyblue2", high = "skyblue4", midpoint = 10000) +
  # theme refer from RPubs
  theme(panel.background = element_rect(fill = "grey"),
        plot.background = element_rect(fill = "grey"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        panel.border =  element_blank(),
        plot.title = element_text(
          size = 15, hjust = 0.5, family = "Times",colour = "white"),
        legend.title = element_text(
          hjust = 0.4 ,vjust = 0.3, size = 10,family = "Times"),
        legend.text = element_text( hjust = 0.4 ,vjust = 2, size = 8,family = "Times"))


```


