---
title: "Personal Consumption Expenditures by State"
author: "Jibei Zheng"
date: "11/18/2021"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---


```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(usmap)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%"
)
```


```{r, echo = FALSE, message = FALSE}
raw_df = 
  read_excel("data/pce_by_state.xlsx", sheet = "Table 1", range = "A4:F63", col_names = FALSE)

colnames(raw_df) = c("state", "pce_2018", "pce_2019", "pce_2020", "change_2019", "change_2020")

pce_state_df_1 = 
  raw_df %>% 
  filter(!(state %in% c("United States", "New England", "Mideast", "Great Lakes", "Plains", "Southeast", "Southwest", "Rocky Mountain", "Far West"))) %>% 
  mutate(
    code = c("CT", "ME", "MA", "NH", "RI", "VT", "DE", "DC", "MD", "NJ", "NY", "PA", "IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD", "AL", "AR", "FL", "GA", "KY", "LA", "MS", "NC", "SC", "TN", "VA", "WV", "AZ", "NM", "OK", "TX", "CO", "ID", "MT", "UT", "WY", "AK", "CA", "HI", "NV", "OR", "WA")
  ) %>% 
  mutate(
    text_2020 = paste(state, "\n", "PCE:", pce_2020),
    text_2019 = paste(state, "\n", "PCE:", pce_2019),
    text_change_2020 = paste(state, "\n", "Percent change:", change_2020),
    text_change_2019 = paste(state, "\n", "Percent change:", change_2019)
  )

personal_df = 
  read_excel("data/pce_by_state.xlsx", sheet = "Table 4", range = "A4:G63", col_names = FALSE)

colnames(personal_df) = c("state", "personal_pce_2020", "housing", "health", "food", "gasoline", "other")

pce_state_df_1 = 
  left_join(pce_state_df_1, personal_df, by = "state") %>% 
  mutate(
    population = pce_2020 / personal_pce_2020,
    personal_pce_2019 = round(pce_2019 / population),
    text_personal_2020 = paste(state, "\n", "Personal PCE:", personal_pce_2020),
    text_personal_2019 = paste(state, "\n", "Personal PCE:", personal_pce_2019)
  )
```

```{r, echo = FALSE, message=FALSE}
# load data set of covid by states
postal_code = 
  read_csv("./data/us_postal_code.csv") %>% 
  janitor::clean_names()

read_data_function = function(file_df) {
  
  state_df = 
    read_csv(file_df, skip = 2)
  
  return(state_df)
}
# create a data frame containing all participants
state_2020 =
  tibble(
    files = list.files("./data/state_covid_separate")
  ) %>% 
  mutate(
    path = map(.x = files, ~paste("./data/state_covid_separate", ., sep = "/"))
  ) %>%                                                             # add path
  mutate(
    observations = map(path, read_data_function))

state_2020_tidy =
  state_2020 %>% 
  unnest(cols = "observations") %>% 
  janitor::clean_names() %>% 
  filter(date == "Dec 31 2020") %>%
  select(state, total_cases, total_case_rate_per_100k)

covid_state_df = left_join(state_2020_tidy, postal_code, by = "state") %>%
  drop_na()
```


## Consumption and Covid by State

```{r, message=FALSE, warning=FALSE}
# create joint data frame
jointly_df = 
  left_join(pce_state_df_1, covid_state_df, by = c("state", "code")) %>% 
  mutate(state = as.factor(state)) %>% 
  select(state, code, pce_2020, total_case_rate_per_100k, everything())

# make plot
Con_case_by_state =
jointly_df %>% 
  arrange(total_case_rate_per_100k) %>% 
  mutate(state = fct_reorder(state, total_case_rate_per_100k)) %>% 
  plot_ly(x = ~state) %>% 
  add_trace(y = ~pce_2020, type = "scatter", mode = "lines",name = "Consumption", yaixs = "y") %>% 
  add_trace(y = ~total_case_rate_per_100k, type = "bar", name = "Covid Cases", yaxis = "y2", opacity = 0.6, marker = list(color = 'rgb(158,202,225)'))  %>% 
  layout(title = "Consumption of 2020 with Covid Cases",
         yaxis = list(title = "consumption expenditure", side = "left"),
         yaxis2 = list(title = "covid cases rate", side = "right", overlaying = "y"),
         showlegend = TRUE)

Con_case_by_state
```









State personal consumption expenditures (PCE) decreased 2.6 percent in 2020 after increasing 3.7 percent in 2019, according to statistics released today by the U.S. Bureau of Economic Analysis (BEA). The percent change in PCE across all 50 states and the District of Columbia ranged from 1.2 percent in Idaho and Utah to –5.8 percent in the District of Columbia. PCE increased in only four states, Idaho, Utah, Montana, and Arkansas.

```{r, warning = FALSE}
g = list(
  scope = 'usa',
  projection = list(type = 'albers usa')
)

fig_1 = 
  plot_geo(pce_state_df_1, locationmode = 'USA-states') %>% 
  add_trace(
    type = "scattergeo",
    locations = ~code,
    text = ~code,
    mode = "text",
    textfont = list(color = rgb(0,0,0), size = 10),
    hoverinfo = "none"
  ) %>% 
  add_trace(
    z = ~change_2020, 
    locations = ~code,
    text = ~text_change_2020,
    color = ~change_2020, 
    colorscale = list(c(0, 0.85, 1), c("#c44934", "#ffffff", "#83c5fd")),
    hoverinfo = "text",
    colorbar = list(title = "PCE growth rates(%)", thickness = 20, x = 1, y = 0.8),
    visible = T
  ) %>% 
  add_trace(
    z = ~change_2019, 
    locations = ~code,
    text = ~text_change_2019,
    color = ~change_2019, 
    colorscale = list(c(0, 1), c("#f5faff", "#0863b0")),
    hoverinfo = "text",
    colorbar = list(title = "PCE growth rates(%)", thickness = 20, x = 1, y = 0.8),
    visible = F
  ) %>%
  layout(
    title = 'Personal Consumption Expenditures by State: Percent Change',
    geo = g,
    updatemenus = list(
      list(
        type = 'buttons',
        x = 0.1,
        y = 0.95,
        buttons = list(
          list(method = "restyle",
               args = list("visible", list(T, T, F)),
               label = '2019-2020'),
          
          list(method = "restyle",
               args = list("visible", list(T, F, T)),
               label = '2018-2019')
        )))
  )

fig_1
```

Across all states and the District of Columbia, per capita PCE was $42,635. Per capita PCE by state ranged from a high of $52,001 in Massachusetts to a low of $32,358 in Mississippi. Per capita PCE in the District of Columbia was $65,169. 

```{r}
fig_3 = 
  plot_geo(pce_state_df_1, locationmode = 'USA-states') %>% 
  add_trace(
    type = "scattergeo",
    locations = ~code,
    text = ~code,
    mode = "text",
    textfont = list(color = rgb(0,0,0), size = 10),
    hoverinfo = "none"
  ) %>% 
  add_trace(
    z = ~personal_pce_2020, 
    locations = ~code,
    text = ~text_personal_2020,
    color = ~personal_pce_2020, 
    colorscale = list(c(0, 0.3, 0.5, 1), c("#f5faff", "72b5f1", "176fba", "#004c8c")),
    hoverinfo = "text",
    colorbar = list(title = "PCE", thickness = 20, x = 1, y = 0.8),
    visible = T
  ) %>% 
  add_trace(
    z = ~personal_pce_2019, 
    locations = ~code,
    text = ~text_personal_2019,
    color = ~personal_pce_2019, 
    colorscale = list(c(0, 0.26, 0.48, 1), c("#f5faff", "72b5f1", "176fba", "#004c8c")),
    hoverinfo = "text",
    colorbar = list(title = "PCE", thickness = 20, x = 1, y = 0.8),
    visible = F
  ) %>% 
  layout(
    title = 'Per Capita Personal Consumption Expenditures by State',
    geo = g,
    updatemenus = list(
      list(
        type = 'buttons',
        x = 0.1,
        y = 0.95,
        buttons = list(
          list(method = "restyle",
               args = list("visible", list(T, T, F)),
               label = '2020'),
          
          list(method = "restyle",
               args = list("visible", list(T, F, T)),
               label = '2019')
        ))),
    annotations = list(list(text = "Year", x = 0, y = 1, showarrow = FALSE))
  )

fig_3
```

```{r}
fig_2 = 
  plot_geo(pce_state_df_1, locationmode = 'USA-states') %>% 
  add_trace(
    type = "scattergeo",
    locations = ~code,
    text = ~code,
    mode = "text",
    textfont = list(color = rgb(0,0,0), size = 10),
    hoverinfo = "none"
  ) %>% 
  add_trace(
    z = ~pce_2020, 
    locations = ~code,
    text = ~text_2020,
    color = ~pce_2020, 
    colorscale = list(c(0, 0.1, 0.5, 1), c("#f5faff", "72b5f1", "176fba", "#005196")),
    hoverinfo = "text",
    colorbar = list(title = "PCE", thickness = 20, x = 1, y = 0.8),
    visible = T
  ) %>% 
  add_trace(
    z = ~pce_2019, 
    locations = ~code,
    text = ~text_2019,
    color = ~pce_2019, 
    colorscale = list(c(0, 0.09, 0.43, 1), c("#f5faff", "72b5f1", "176fba", "#004c8c")),
    hoverinfo = "text",
    colorbar = list(title = "PCE", thickness = 20, x = 1, y = 0.8),
    visible = F
  ) %>% 
  layout(
    title = 'Total Personal Consumption Expenditures by State',
    geo = g,
    updatemenus = list(
      list(
        type = 'buttons',
        x = 0.1,
        y = 0.95,
        buttons = list(
          list(method = "restyle",
               args = list("visible", list(T, T, F)),
               label = '2020'),
          
          list(method = "restyle",
               args = list("visible", list(T, F, T)),
               label = '2019')
        ))),
    annotations = list(list(text = "Year", x = 0, y = 1, showarrow = FALSE))
  )

fig_2
```

Expenditures on food services and accommodations, recreation services, and health care were the leading contributors to the decrease nationally.

Food services and accommodations decreased 20.5 percent nationally and was the largest contributor to decreases in the District of Columbia and 27 states including Hawaii and New York; it was the second largest contributor to decreases in 11 states.

Recreation services decreased 28.6 percent nationally and was the largest contributor to decreases in 4 states including California.

Health care decreased 6.1 percent nationally and was the largest contributor to decreases in 12 states including Alaska.


```{r, include = FALSE}
## us = usmap::us_map()
## 
## us_val =
##   us %>% 
##   left_join(pce_state_20_df, by = c("full" = "state"))
## 
## us_centroids = 
##   us_val %>% 
##   group_by(full) %>% 
##   summarise(centroid.x = mean(range(x)), 
##             centroid.y = mean(range(y)),
##             label = abbr,
##             pce = pce)
## 
## state_plot =
##   us_val %>% 
##   ggplot() +
##   geom_polygon(
##     aes(x,y, group = group, fill = pce), 
##     color = "black",
##     size = .1
##   ) +
##   geom_text(
##     data = us_centroids, 
##     aes(centroid.x, centroid.y, label = paste(label, "\n", pce)),
##     size = 2
##   ) +
##   scale_fill_gradient2(low = "white", mid = "skyblue2", high = "skyblue4", midpoint = 800000) +
##   labs(
##     title = "Per Capita Personal Consumption Expenditures by State, 2020",
##     x = "",
##     y = ""
##   ) +
##   theme_void()
## 
## ggplotly(state_plot)
```

